Malacology: A Programmable Storage System

### 2016-10-06 paper/paper.tex
paper: first crack at decoupling Ceph from intro - did a hack job on the contributions (more general) - adopted the term underlying storage instead of the word Ceph - split the intro into 2 section to shore up motivation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{abstract}
Storage systems are caught between rapidly changing data processing systems and
the increasing speed of storage devices. This puts tremendous pressure on
storage systems to adapt both in terms of their interfaces and their
performance. But adapting storage systems can be difficult because unprincipled
changes might jeopardize years of code-hardening and performance optimization
efforts that were necessary for users to entrust their data to the storage
system.  We introduce  Malacology, a prototype programmable storage system to
explore how existing abstractions of common services found in storage systems
can be leveraged to address new data processing systems and the increasing
speed of storage devices. This approach allows unprecedented flexibility for
storage systems to evolve without sacrificing the robustness of its
code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by constructing two services out of existing abstractions: a
file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to make trade-offs for lease capabilities, balance load
(2\(\times\) improvement), propagate interfaces (average of 222ms) and recover
(15 second down time).  
\end{abstract}

@@

A number of recent developments are disrupting traditional storage systems: (1)
the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}; (2) demand for managing
structured data and flexible consistency semantics at scale pressure big data
processing systems to use storage abstractions that can meet these
demands~\cite{apache_contributors_parquet_2014}; and (3) production-quality
scalable storage systems available as open source software have established and
are continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

In this paper we investigate an approach that focuses on generalizing
existing storage system resources, services, and abstractions that in
generalized form can be used to \emph{program} new services. By doing so
one can reuse subsystems and leverage their optimizations,
established correctness, robustness, and efficiency. We will refer to
this programmability as \emph{programmable storage}, which differs from
\emph{active storage} (the injection and execution of arbitrary code in
a storage system) and \emph{software-defined storage} (the control of
thin-provisioning of storage).

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system for constructing new
services by re-purposing existing subsystem abstractions of the storage stack.
We build the framework in Ceph by leveraging a broad spectrum of existing
services, including distributed locking and caching services provided by
metadata servers, durability and logical storage devices provided by the
backend object store, and propagation of consistent cluster state provided by
the monitoring service (see Figure~\ref{fig:overview}). As we will show in this
paper, this framework is expressive enough to provide the functionality
necessary for implementing new services. Our contributions are:

\paragraph*{A programmable storage system} that re-uses and extends existing
abstractions and services of Ceph, a common production-level storage system. It
includes:

@@

\item An extension of file system metadata by a new type of ``sequencer''
inode that together with Ceph's capability subsystem implements
high-performance, non-durable access serialization:

\item The use of Ceph's PAXOS-based consensus infrastructure for versioning
of metadata service load-balancing policies

\item The use of storage objects for persisting policies

\item An extension of the object interface classes for new logical
storage/metadata devices

@@

\paragraph*{A distributed shared log service} to demonstrate the feasibility of
services using the programmable storage approach:

@@

\item An application of Mantle to manage load-balancing of sequencers.

@@

The remainder of this paper is structured as follows. First we describe
Malacology by presenting the subsystems within Ceph that we re-purpose, and
briefly describe how those system are used within Malacology
(\S\ref{sec:malacology}). Next we describe the services that we have
constructed within the Malacology framework (\S\ref{sec:services}), and
evaluate our ideas within our prototype implementation (\S\ref{evaluation}).
Finally we discuss related work and conclude.

@@

To illustrate the benefits and challenges of this approach we examine the

Today's storage stacks are large, monolothic systems built to ensure
scalability and persistence -- a storage system that loses user's data is
useless. Building storage stacks from the ground up for a specific purpose
results in the best performance. For example, HDFS was designed specifically to
server Hadoop's jobs and relaxes POSIX constraints~\cite{CITEME}.
Alternatively, general-purpose storage stacks are built with the flexiblity to
serve many applications by embedding tunable parameters to control low level
implementation details like block size. Unfortunately, users want more control
from the general-purpose storage stacks without going as far as building their
storage system from the ground up.

% does this belong?
To demonstarte the need for programmability in storage systems, we examine
Ceph's programmability. Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
popular, production-level open-source distributed storage system, was
originally built to be an HPC file system. But demands for storage systems that
could slide in behind cloud infrastructures, like OpenStack, shifted the focus
to Ceph's internal object store as a back end for large cloud block devices,
S3-like objects, and file system workloads.  Something of a storage swiss army
knife, Ceph supports file, block, and object interfaces simultaneously in a
single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of storage devices
(OSDs) that provide Ceph with data durability and integrity using replication,
erasure-coding, and scrubbing~\cite{weil_rados_2007}. 
% By introducing programmability concepts into Ceph, we can build new services
% by carefully exposing internal storage services to applications.

Ceph also provides empirical evidence that developers embrace programmable
storage: Ceph already supports the implementation of domain-specific logical
object interfaces by composing existing interfaces. These interfaces are
written in C++ and are statically loaded into the system.
Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production
use of domain-specific interfaces in the Ceph community since 2010.  What is
most remarkable is that this trend contradicts the notion that API changes are
a burden for users.  Rather it appears that a gap in existing interfaces are
being addressed through programmability. In fact, Table~\ref{table:objclasses}
categorizes existing interfaces and we clearly see a trend towards reusable
services.

@@

The popularity of the custom object interface facility of Ceph hint at three
trends in the Ceph community: (1) increasingly, the default algorithms/tunables
of the storage system are insufficient for the application's performance goals,
(2) programmers are becoming more aware of their application's behavior, and
(3) programmers know how to manage resources to improve performance.
Programmers gravitate towards object interfaces because it gives them the
ability to tell the storage system about their application: if it is CPU or IO
bound, if it has locality, if its size has the potential to overload a single
proxy node, etc. The programmers know what the problem is and how to solve it,
but until the ability to modify object interfaces, they had no way to express
to the storage system how to handle their data.

@@

\label{services}

\begin{table}
\centering
\begin{tabular}{  l | l | l    }
\textbf{Service} & \textbf{Ceph Component} & \textbf{Used for...}  \\ \hline
Mantle  & RADOS    & policy durability \\
        & MON maps & policy versioning \\ \hline
ZLog    & Logical Storage & CORFU protocol  \\ 
        & MON maps & interface install \\ \hline
Seq.    & MDS file types & log management \\ 
        & MDS fault toler. & seq. fault toler. \\
\end{tabular}
\caption{To implement the Mantle and ZLog services, we leverage Ceph
components. Malacology uses these subsystems because they provide powerful
abstractions.}
\label{table:implementation}
\end{table}

>>>>

\begin{abstract}
Storage systems are caught between rapidly changing data
processing systems and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to adapt both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows unprecedented
flexibility for storage systems to evolve without sacrificing the robustness of
its code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by constructing two services out of existing abstractions: a
file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to make trade-offs for lease capabilities, balance load, propogate
interfaces, and recover from failure.
\end{abstract}

@@

A number of recent developments are disrupting traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}. Second, the demand for
managing structured data and flexible consistency semantics at scale pressure
big data processing systems to use storage abstractions that can meet these
demands~\cite{apache_contributors_parquet_2014}. Finally, production-quality
scalable storage systems available as open source software have established and
are continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

In this paper we investigate an approach that focuses on generalizing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By doing so one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency. We will refer to this programmability as
\emph{programmable storage}, which differs from \emph{active storage} (the
injection and execution of arbitrary code in a storage system) and
\emph{software-defined storage} (the control of thin-provisioning of storage).

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system for constructing new
services by re-purposing existing subsystem abstractions of the storage stack.
We build the framework in Ceph, a popular, open source software storage stack,
by leveraging a broad spectrum of existing services, including distributed
locking and caching services provided by metadata servers, durability and
logical storage devices provided by the backend object store, and propagation
of consistent cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, this framework is
expressive enough to provide the functionality necessary for implementing new
services. Our contributions are:

\paragraph*{A programmable storage system} that re-uses and extends existing
abstractions and services of the underlying storage stack. It
includes the following building blocks:

@@

\item Versioning interface that uses the underyling storage stack's consensus
infrastructure

\item Logical Device interface for adding functionality to daemons that manage
storage/metadata devices, allowing users to push computation to the data

\item An interface for mediating access to a shared resources, which exposes
serialization primatives using capabilities, caching, and batching

\item Load Balancing interface for managing resources across the cluster in
multi-tenant environments

\item Durability interface for persisting policies using the underyling storage
stack's object store

@@

\paragraph*{Two distributed services} that demonstrate the feasibility of using
the programmable storage approach:

@@

@@

The remainder of this paper is structured as follows. First, we describe and
motivate the need for programmable storage by describing current practices in
the open source community. Next we describe Malacology by presenting the
subsystems within the underlying storage system that we re-purpose, and briefly
describe how those system are used within Malacology (\S\ref{sec:malacology}).
Then we describe the services that we have constructed within the Malacology
framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
implementation (\S\ref{evaluation}).  Finally we discuss related work and
conclude.

@@

Building storage stacks from the ground-up for a specific purpose results in
the best performance. For example, HDFS was designed specifically to serve
Hadoop's jobs, and uses techniques like relaxing POSIX constraints to achieve
the best performance~\cite{CITEME}. Alternatively, general-purpose storage
stacks are built with the flexiblity to serve many applications by exposing
tunable parameters to control low level implementation details like block size.
Unfortunately, users want more control from the general-purpose storage stacks
without going as far as building their storage system from the ground up.

% does this belong?
%Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
%popular, production-level open-source distributed storage system, was
%originally built to be an HPC file system. But demands for storage systems that
%could slide in behind cloud infrastructures, like OpenStack, shifted the focus
%to Ceph's internal object store as a backend for large cloud block devices,
%S3-like objects, and file system workloads.

To demonstrate this trend in storage systems, we examine Ceph's
programmability. Something of a storage swiss army knife, Ceph supports file,
block, and object interfaces simultaneously in a single
cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of storage devices
(OSDs) that provide Ceph with data durability and integrity using replication,
erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already has some
programmability; the OSDs support the implementation of domain-specific logical
object interfaces by composing existing interfaces. These interfaces are
written in C++ and are statically loaded into the system.

% By introducing programmability concepts into Ceph, we can build new services
% by carefully exposing internal storage services to applications.

The Ceph community provides empirical evidence that developers embrace
programmable storage. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic
growth in the production use of domain-specific interfaces in the Ceph
community since 2010.  What is most remarkable is that this trend contradicts
the notion that API changes are a burden for users.  Rather it appears that a
gap in existing interfaces are being addressed through programmability. In
fact, Table~\ref{table:objclasses} categorizes existing interfaces and we
clearly see a trend towards reusable services.

@@

The takeway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
already trying to achieve programmability because their needs, whether they be
related to performance, availability, consistency, correctness etc., are not
satisfied by the underlying storage system. The popularity of the custom object
interface facility of Ceph could be due to a number of reasons, such as the
default algorithms/tunables of the storage system are insufficient for the
application's performance goals, programmers want to exploit
application-specific semantics, and/or programmers know how to manage resources
to improve performance. Their solution is a way to work around the
traditionally rigid storage APIs because custome object interfaces give
programmers the ability to tell the storage system about their application: if
it is CPU or IO bound, if it has locality, if its size has the potential to
overload a single proxy node, etc.  The programmers know what the problem is
and how to solve it, but until the ability to modify object interfaces, they
had no way to express to the storage system how to handle their data.

Our approach is to expose more of the commonly used, code-hardened subsystems
of the underlying storage system as interfaces. The intent is that these
interfaces, which can be as simple as a redirection to the persistent data
store or as complicated as a strongly consistent directly service, should be
used and re-used in many contexts to implement a wide range of services. By
making programmability a `feature', rather than a `hack' or `workaround', we
help standardize a development process that is largely ad-hoc.

@@

\label{services}
\begin{table}
\centering
\begin{tabular}{  l | l | l | l }
\textbf{Service} & \text{*} & \textbf{Malacology Collection of Interfaces/APIS}       & \textbf{Used to...}  \\ \hline
Mantle           &          & Load Balancing      & load balance metadata \\
                 & \cmark   & Versioning          & agree on current balancer \\ 
                 & \cmark   & Batching            & control client write lock \\
                 &          & Durability          & persist policy logic \\\hline
ZLog             & \cmark   & Logical Device      & implement CORFU \\ 
                 & \cmark   & Versioning          & install ZLog interfaces \\
                 &          & Durability          & persist log entries \\ \hline
ZLog             &          & Load balancing      & load balance sequencers \\
Seqr.            & \cmark   & Batching            & guard shared resource \\ %\hline
                 %\multicolumn{4}{c}{} \\ 
                 %\multicolumn{4}{c}{\#: interface in metadata server,     \(\mp\): interface in object storage server} \\ 
\end{tabular}
\caption{Mantle and ZLog are built using Malacology hooks so they inherit the
robustness of the subsystems of the underlying stack. For the \text{*} column,
the \cmark s are hooks contributed in this paper; durability is already
provided with a client library and the MDS Logical Device is already
published~\cite{sevilla:sc15-mantle}. }
\label{table:implementation}
\end{table}
